---
title: "`r params$report_title`"
author: "`r params$report_author`"
date: "`r params$report_time`"
output:
  html_document:
    toc: True
params:
  report_title: "Show transcript feature types for a single sample"
  report_author: "Anonymus"
  report_time: !r format(Sys.Date(), format="%d/%m/%Y")
  input_format: "pseudo_tab"
  input_path: "/home/rstudio/local_files/example_data/pseudo_tabular_format.csv"
  output_folder: "."
---

## Setup

```{r}
library(dplyr)
library(tidyr)
library(tibble)
```

## Parse input

```{r}
in_tab <- read.csv(params$input_path)
```

### Reshape if pseudotabular

Data might come as a pseudo-table with entities in column names. This has to be
as a minimum, reshaped to long format.

```{r}
if (params$input_format == "pseudo_tab") {
  main_tab <- in_tab %>%
    pivot_longer(everything()) %>%
    rename(Entity = name, Label = value) %>%
    dplyr::filter(!is.na(Label))
}
```

### Reshape if nested list

Data might come as list of lists, so those lists have to be unnested.

```{r}
if (params$input_format == "nested") {
  in_tab[[2]] <- strsplit(in_tab[[2]], ";")
  colnames(in_tab <- c("Entity", "Label"))
  main_tab <- unnest(in_tab, Label)
}
```

### Pivot wider after coercing to long format

Long format data has to cast into a one-hot encoded matrix..

```{r}
if (params$input_format == "long") {
  main_tab <- in_tab
}
```

```{r}
if (params$input_format %in% c("long", "nested", "pseudo_tab")) {
  main_matrix <- main_tab %>%
    mutate(value = 1) %>%
    pivot_wider(names_from=Label, values_fill=0) %>%
    column_to_rownames("Entity")
} else {
  main_matrix <- column_to_rownames(in_tab, "Entity")
}

main_matrix[seq(1, 5), seq(1, 5)]
```

## Create distance matrix

Under the present circumstances, the ***binary*** method of the base-R dist function
is equivalent to creating Jaccard's distances.

```{r}
distmat <- dist(main_matrix, method="binary")

dist_tab <- distmat %>%
  {1 - .} %>%
  as.matrix() %>%
  data.frame() %>%
  rownames_to_column("Entity")

dist_tab[seq(1, 5), seq(1, 5)]
```

This can directly be saved after converting it to a table

```{r}
write.csv(
  dist_tab, file.path(params$output_folder, "distance_matrix.csv"),
  row.names=FALSE
)
```

Perhaps a better means of data representation is a long format

```{r}
dist_tab %>%
  pivot_longer(-Entity, values_to="Similarity", names_to="Pair") %>%
  filter(Entity != Pair) %>%
  write.csv(
    file.path(params$output_folder, "distance_matrix.csv"), row.names=FALSE
  )
```

## Visualization options
