---
title: "`r params$report_title`"
author: "`r params$report_author`"
date: "`r params$report_time`"
output:
  html_document:
    toc: True
params:
  report_title: "Generate example data"
  report_author: "Anonymus"
  report_time: !r format(Sys.Date(), format="%d/%m/%Y")
  output_folder: "/home/rstudio/local_files/example_data"
---

## Setup

```{r}
library(dplyr)
library(tidyr)
library(tibble)
```

## Generate toy data

### One-hot-encoded format

To generate a dataset with close to realistic dimensions, 20 000 entities are
linked to 5000 distinct labels randomly.  
Labels will be one-hot-encoded, where 1 in the matrix means the labels is associated
with the entity and 0 means it is not.

```{r}
example_df <- data.frame(Entity = paste("e", seq(1001, 3000), sep="_"))

for(cl in paste("cat", seq(1001, 1050), sep="_")){
  example_df[[cl]] <- sample(c(0, 1), 2000, replace=TRUE)
}
```

Save the generated toy data in this native format:

```{r}
write.csv(
  example_df, file.path(params$output_folder, "onehot_encoded_matrix.csv"),
  row.names=FALSE
)
```

### Conversion to long format

The large number of columns is not optimal for R to handle data. One possible way
to make the data tidier is to reshape it into long format.

```{r}
long_df <- pivot_longer(example_df, -Entity)
```

This way it is easier to show what we have in our dataframe:

```{r}
head(long_df)
```






